## R Markdown

Spearman Rank Correlation between reponse varaibles: ObservationCount1 (2 min silent listening period) and ObservationCount (total count).

Install the required packages 
```{r install packages, echo=FALSE, message=FALSE}

library(tidyverse)
library(reshape)
library(Hmisc)

```

Code to read and append the relevant files from the out directory. 
Note: I manaully added the SpeicesCode to these files so that they could be distinguided. In future iterations, this should be added to the tables druing the analysis. 

```{r pull multiple files, echo=FALSE, message=FALSE}

#Read in the events data then filer for just ATOWLS 
require(plyr)

out.dir<-"output"
my.files<-list.files(path = out.dir, patter ="*PosteriorSummary.csv", full.name=TRUE)
res1<-ldply(my.files, read.csv)
res1$Response<-"silent"

my.files<-list.files(path = out.dir, patter ="*PosteriorSummary2.csv", full.name=TRUE)
res2<-ldply(my.files, read.csv)
res2$Response<-"total"

#bind the two table together 
res<-rbind(res1, res2)

detach(package:plyr) #detach or this will cause an issue with dplyr

```

Write a speices specific loop to do the Spearman rank correlation. This will be done on both the abundance and trend paramaters and for each speices seperatly (with the exception of BOOW, which only have data for Response=="total")

Since each province has the same playback protocol (mostly) in the Altantic region, these data can be pooled. If comparing across regions with different protocols, then they cannot necassrily be pooled. 


```{r SpearmanTau, echo=FALSE, message=FALSE}

res<-res %>% filter(Species!="Boreal Owl")

#prepare dataframe

tau<-cast(res, Species+StateProvince+grid_id~Response, value="tau")
tau<-tau %>% rename(silent_tau = silent, total_tau=total)

#to plot upper and lower CI
tau_cl<-cast(res, Species+StateProvince+grid_id~Response, value="tau_ll")
tau_cl<-tau_cl %>% select(silent, total) %>% rename(silent_tau_cl = silent, total_tau_cl=total)

tau_cu<-cast(res, Species+StateProvince+grid_id~Response, value="tau_ul")
tau_cu<-tau_cu %>% select(silent, total) %>% rename(silent_tau_cu = silent, total_tau_cu=total)

tau<-cbind(tau, tau_cl, tau_cu)

sp.list<-unique(res$Species)

for (i in 1:length(sp.list)){
  
  i<-3 #for testing
  
  sp.tau<-tau %>% filter(Species==sp.list[i])
  
  sp.tau<-na.omit(sp.tau)
  
  cor.test(sp.tau$silent_tau, sp.tau$total_tau, method = "spearman")

#plot relationship  
ggplot(data=sp.tau, aes(silent_tau, total_tau))+
  geom_point()+
#geom_errorbar(aes(ymin = total_tau_cl,ymax = total_tau_cu)) + 
#geom_errorbarh(aes(xmin = silent_tau_cl,xmax = silent_tau_cu))+
  geom_abline()+
  geom_hline(yintercept=0, linetype="dotted")+
  geom_vline(xintercept=0, linetype="dotted")+
  labs(x="Silent Listening Period Count", y="Total Count")
  

ggplot(data = df,aes(x = x,y = y)) + 
    geom_point() + 
    geom_errorbar(aes(ymin = ymin,ymax = ymax)) + 
    geom_errorbarh(aes(xmin = xmin,xmax = xmax))

  
  
}# end speices loop


```

```{r SpearmanAlph, echo=FALSE, message=FALSE}

res<-res %>% filter(Species!="Boreal Owl")

alph<-cast(res, Species+StateProvince+grid_id~Response, value="alph")
alph<-alph %>% rename(silent_alph = silent, total_alph=total)

#to plot upper and lower CI
alph_cl<-cast(res, Species+StateProvince+grid_id~Response, value="alph_ll")
alph_cl<-alph_cl %>% select(silent, total) %>% rename(silent_alph_cl = silent, total_alph_cl=total)

alph_cu<-cast(res, Species+StateProvince+grid_id~Response, value="alph_ul")
alph_cu<-alph_cu %>% select(silent, total) %>% rename(silent_alph_cu = silent, total_alph_cu=total)

alph<-cbind(alph, alph_cl, alph_cu)

sp.list<-unique(res$Species)

for (m in 1:length(sp.list)){
  
  m<-3 #for testing
  
  sp.alph<-alph %>% filter(Species==sp.list[m])
  
  sp.alph<-na.omit(sp.alph)
  
  cor.test(sp.alph$silent_alph, sp.alph$total_alph, method = "spearman")

#plot relationship  
ggplot(data=sp.alph, aes(silent_alph, total_alph))+
  geom_point()+
  #geom_abline()+
  #geom_hline(yintercept=0, linetype="dotted")+
  #geom_vline(xintercept=0, linetype="dotted")+
  #geom_errorbar(aes(ymin = total_alph_cl,ymax = total_alph_cu)) + 
  #geom_errorbarh(aes(xmin = silent_alph_cl,xmax = silent_alph_cu))+
  labs(x="Silent Listening Period Count", y="Total Count")
  
  
  
}# end speices loop


```

Compare CI widths to assess the amount the precision of trend estiamtes

```{r CI Width, echo=FALSE, message=FALSE}

res<-res %>% filter(Species!="BOOW")

sp.list<-unique(res$Species)

for (k in 1:length(sp.list)){
  
  k<-1 #for testing
  
  sp.ci<-res %>% filter(Species==sp.list[k])
  
  ggplot(data=sp.ci, aes(Response, tau_iw), group=Response)+
  geom_boxplot()
  
#paired t-test to look at difference before and after
  
  cast.ci<-cast(sp.ci, Species+StateProvince+grid_id~Response, value="tau_iw")
  cast.ci<-na.omit(cast.ci)
  
cast.ci$difference <- cast.ci$total - cast.ci$silent

ggplot(cast.ci) +
  aes(y = difference) +
  geom_boxplot() +
  theme_minimal()
  
test <- t.test(cast.ci$total, cast.ci$silent,
  alternative = "less",
  paired = TRUE)

test  
test$p.value

}

```

Prepare data for plotting in ArcGIS

```{r cast multiple files, echo=FALSE, message=FALSE}

res<-res %>% filter(Species!="Boreal Owl")

out.dir<-"output/2019"

#recast tables
alph<-cast(res, Species+StateProvince+grid_id~Response, value="alph")
alph<-alph %>% rename(silent_alph = silent, total_alph=total)

#to plot upper and lower CI
alph_cl<-cast(res, Species+StateProvince+grid_id~Response, value="alph_ll")
alph_cl<-alph_cl %>% select(silent, total) %>% rename(silent_alph_cl = silent, total_alph_cl=total)

alph_cu<-cast(res, Species+StateProvince+grid_id~Response, value="alph_ul")
alph_cu<-alph_cu %>% select(silent, total) %>% rename(silent_alph_cu = silent, total_alph_cu=total)

alph<-cbind(alph, alph_cl, alph_cu)

tau<-cast(res, Species+StateProvince+grid_id~Response, value="tau")
tau<-tau %>% rename(silent_tau = silent, total_tau=total)

#to plot upper and lower CI
tau_cl<-cast(res, Species+StateProvince+grid_id~Response, value="tau_ll")
tau_cl<-tau_cl %>% select(silent, total) %>% rename(silent_tau_cl = silent, total_tau_cl=total)

tau_cu<-cast(res, Species+StateProvince+grid_id~Response, value="tau_ul")
tau_cu<-tau_cu %>% select(silent, total) %>% rename(silent_tau_cu = silent, total_tau_cu=total)

tau<-cbind(tau, tau_cl, tau_cu)

post<-full_join(alph, tau, by=c("Species", "StateProvince", "grid_id"))


sp.list<-unique(post$Species)

for (l in 1:length(sp.list)){
  
  #l<-1 #for testing
  
  sp.post<-post %>% filter(Species==sp.list[l])
  
  write.table(sp.post, paste(out.dir, sp.list[l], "PosteriorArcPlot.csv", sep="."), row.names = FALSE, append = FALSE, quote = FALSE, sep = ",", col.names = TRUE)
  
}

```